---
title: "HW2_code"
output: github_document
---

# Problem 6

A study is to examine the associations of depression and cognitive performance with migraine symptoms: 

## Some backgrounds:

CESD: questionairre screening for depression; meausre symptoms of depression; sum of scores to all questions; longer duration of symptoms will have greater score.

nddie: The Neurological Disorders Depression Inventory for Epilepsy (NDDI-E) is a 6-item questionnaire validated to screen for depression in people with epilepsy


## Import data

```{r setup, include=FALSE}
library(tidyverse)
```


```{r import data}
library(readxl)

excel_sheets("Migraine.xlsx")
Mig <- read_excel("Migraine.xlsx")

Mig <- janitor::clean_names(Mig) # rename variable names
```

Have a look at the data:

```{r}
head(Mig)
str(Mig)
```

variables in the data:

| Variable names |   description  |
|:---------------:|:-----------------|
|Migraine        |Migraine status: 0(no), 1(yes) |
|CESD            |Depression Scale|
|NDDIE           |Self rating tool to screen for depression, inventory|
|ABNAS memory    |Assessement for memory |
|ABNAS language  |Assessment for language |

## exploratory analysis

Seperate subjects with and without migraine.

```{r seperate variables}
with_Mig <- filter(Mig, migraine == 1)
no_Mig <- filter(Mig, migraine == 0)
```

Explore subjects with migraine

```{r}
summary(with_Mig$cesd)
summary(no_Mig$cesd)
```


plots:

```{r cesd}
par(mfrow = c(1, 2))
hist(no_Mig$cesd, freq = F, ylim = c(0, 0.1))
hist(with_Mig$cesd, freq = F, ylim = c(0, 0.1))
```

```{r nddie}
par(mfrow = c(1, 2))
hist(no_Mig$nddie, freq = F, ylim = c(0, 0.3))
hist(with_Mig$nddie, freq = F, ylim = c(0, 0.3))
```


```{r abnas}
par(mfrow = c(2, 2))
hist(no_Mig$abnas_memory, freq = F, ylim = c(0, 0.6))
hist(with_Mig$abnas_memory, freq = F, ylim = c(0, 0.6))
hist(no_Mig$abnas_language, freq = F, ylim = c(0, 0.7))
hist(with_Mig$abnas_language, freq = F, ylim = c(0, 0.7))
```


summary table for patients with migraine: 

|         |with_mig_cesd  |no_mig_cesd  |with_mig_nddie |no_mig_nddie |with_mig_memory|no_mig_memory|with_mig_language|no_mig_languge|
|---------|---------------|-------------|---------------|-------------|---------------|-------------|---------------|---------------|
|mean     | 
|sd| 
|median   |